Typed Interactions for NLU in Opal

### 2017-12-13 paper.tex
Updated bib
Moved a figure, small revisions

We have seen a huge growth in NLU applications in recent years, and we will
likely see more. Unfortunately, there has been no serious effort made to make
this power available at the language level, so programmers have been forced to
integrate these technologies into their applications by hand. We propose an
addition to the Opal language that will begin to close this gap.

@@

\begin{figure}
\begin{minted}{haskell}
  data Decl = FreeText String
            | Keywords String
                       [String]
            | Trait String
                    [(String, Ty)]
            | Alias String Ty
            deriving (Show)

@@

\subsection{Programmer Freedom}
When working around unsupported types, there is an obvious tendency to modify
the application types to better fit the capabilities of the system. It turns out
that this occurs in more cases than just using dates and times. For example,
consider another calendar example, similar to the one above.
\begin{minted}{typescript}
keywords WkDay = "Sunday" | ...
alias Abs = {
  day: WkDay
  person: Person
}
keywords RDay = "today" | "tomorrow"
alias Rel = {
  day: RDay
  person: Person
}
trait Intent =
  | <Absolute> Abs
  | <Relative> Rel
\end{minted}
Here, we've done away with the complications of dates and times, and we just
allow a user to specify a day of the week or a ``relative day.'' The interesting
observation here is that this is likely not the way that most TypeScript/Opal
developers would write this type. For example, one reasonable alternative might
express days as a top-level disjunction, and then make the type of intents a
normal record.
\begin{minted}{typescript}
  type Day =
    | { tag: "today" }
    | { tag: "tomorrow" }
    | { tag: "abs", day: WkDay }
  type Intent =
    { day: Day, person: Person }
\end{minted}
This makes sense, since \ts{Absolute} and \ts{Relative} are not actully
different \emph{intents}, per se. We can see that the system has led us to a
different type than we would normally write.

@@

% TODO: Fix bib formatting
\bibliography{paper}{}
\bibliographystyle{plain}

>>>>

We have seen a huge growth in NLU applications in recent years, and we will
likely see more. Unfortunately, there has been no serious effort made to make
this power available at the language level, so programmers have been forced to
integrate these technologies into their applications by hand. We propose an
addition to the Opal language that will begin to close this
gap.\cite{opal-arxiv}

@@

\begin{figure}[H]
\begin{minted}{haskell}
  data Decl = FreeText String
            | Keywords String
                       [String]
            | Trait String
                    [(String, Ty)]
            | Alias String Ty
            deriving (Show)

@@

\subsection{Programmer Freedom}
When working around unsupported types, there is an obvious tendency to modify
the application types to better fit the capabilities of the system. It turns out
that this occurs in more cases than just using dates and times. For example,
consider a different set of types for calendar scheduling.
\begin{minted}{typescript}
keywords WkDay = "Sunday" | ...
alias Abs = {
  day: WkDay
  person: Person
}
keywords RDay = "today" | "tomorrow"
alias Rel = {
  day: RDay
  person: Person
}
trait Intent =
  | <Absolute> Abs
  | <Relative> Rel
\end{minted}
Here, we've done away with the complications of dates and times, and we just
allow a user to specify a day of the week or a ``relative day.'' The interesting
observation here is that this is likely not the way that most TypeScript/Opal
developers would write that type. For example, one reasonable alternative might
express days as a top-level disjunction, and then make the type of intents a
normal record.
\begin{minted}{typescript}
  type Day =
    | { tag: "today" }
    | { tag: "tomorrow" }
    | { tag: "abs", day: WkDay }
  type Intent =
    { day: Day, person: Person }
\end{minted}
This makes sense, since \ts{Absolute} and \ts{Relative} are not actully
different \emph{intents}, per se. We can see that the system has led us to a
different type than we would normally write.

@@

\bibliography{paper}{}
\bibliographystyle{plain}
